Single-cell virus sequencing of influenza variants that trigger innate immunity

### 2018-09-10 paper/paper.tex
cleaned up methods for IFN reporter cells

\figsupp[Validation of IFN reporter cell lines.]
{To validate the IFN reporter cell lines, they were infected at high MOI with the Cantell strain of Sendai virus, which strongly activates IFN expression~\citep{strahle2006sendai}.
Activation of the IFN reporter was then monitored by flow cytometry using the marker indicated at the bottom of each plot (either a fluorescent protein or antibody staining for the cell-surface LNGFR).
Sendai infection efficiently activated the IFN reporter in all cases, with the strongest signal for the IFN-$\lambda$ reporter driving ZsGreen.
}
{\includegraphics[width=0.5\textwidth]{figures/IFN_stochastic/IFN_reporter/Sendai_validation.pdf}}
\label{figsupp:reporter_validation}

@@

\subsection{IFN reporter cell lines}
\jdbcomment{Alistair writes this}
For our type I interferon reporters, A 1kb region upstream of the human IFNB1 gene was cloned into the pHAGE2 lentiviral vector, with a NotI site immediately downstream serving as an artificial Kozak sequence (REF). 
Downstream of this NotI site, each of the following reporter constructs was cloned: mCherry, mNeongreen, and mNeongreen translationally linked to low-affinity nerve growth factor lacking the C-terminal signaling domain (LNGFR$\Delta$C) by a p2A linker (REFS).
For our type III interferon reporters, a 1.2kb region upstream of the human IL29 (IFNL1) gene was cloned into the pHAGE2 vector, with the native kozak sequence left at the 3' end. 
Downstream of this promoter zsGreen translationally linked to LNGFR$\Delta$C via a p2A linker was cloned. 
Replication-incompetent lentivirus was generated from these vectors as previously described (REF).
A monolayer of A549 cells in D10 was transduced with each lentivirus in the presence of 5 $\mu$g polybrene, and sorted as single cells. 
These single cell clones were grown and tested for reporter activity by poly I:C transfection---a potent agonist of the RIG-I pathway.
Critically, cells propagated and used in experiments were not exposed to poly I:C,  thus avoiding selection on survival post interferon-induction.
For our dual- type I and type III reporter, a single-cell clone of our type III reporter strain was transduced with our type I reporter bearing the mCherry fluorescent marker and subsequently propagated as a single cell clone and validated as above.

Reporter sequences \FIGDATA[IFNrare]{reporter_sequences}

\subsection{Viral stocks}
\jdbcomment{Alistair writes this}
All viruses were generated using reverse-genetics as previously described.
All infections were performed in influenza growth media as previously described.
Briefly, bidirectional plasmids producing viral mRNA and vRNA are transfected into a co-culture of MDCK-SIAT1 and 293t cells grown in D10.
After 24h, media is changed from D10 to influenza growth media.
Between 48h and 72h post-infection, depending on viral growth kinetics, supernatant was harvested and titered by TCID50.
Virus was subsequently expanded on MDCK-SIAT1 cells as described below.
For all viral variants, the described mutations were introduced into reverse-genetics plasmids using PCR mutagenesis. 
For the HA-flanked mCherry pseudovirus, a plasmid was generated similarly to a previously-described HA-flanked eGFP vector.
Synonymously-barcoded virus utilized previously described 3' synonymous mutations, and two newly-introduced single-nucleotide synonymous mutations at the 5' end of each viral gene.
Wild-type, synonymously-barcoded, NS1-inactivated, and single-nucleotide variant viruses were all grown in unmodified MDCK-SIAT1 and 293t cells.
Viral variants bearing large internal deleltions in the PB1 segment were grown in MDCK-SIAT1 and 293t cells expressing PB1 as previously described.
Pseudovirus expressing mCherry in place of eGFP was grown in MDCK-SIAT1 and 293t cells expressing HA as previously described.

@@

>>>>

\figsupp[Validation of IFN reporter cell lines.]
{To validate the IFN reporter cell lines, they were infected at high MOI with the Cantell strain of Sendai virus, which strongly activates IFN expression~\citep{strahle2006sendai}.
At \jdbcomment{?} hours post-infection, activation of the IFN reporter was then monitored by flow cytometry using the marker indicated at the bottom of each plot (either a fluorescent protein or antibody staining for the cell-surface LNGFR).
Sendai infection efficiently activated the IFN reporter in all cases, with the strongest signal for the IFN-$\lambda$ reporter driving ZsGreen.
}
{\includegraphics[width=0.5\textwidth]{figures/IFN_stochastic/IFN_reporter/Sendai_validation.pdf}}
\label{figsupp:reporter_validation}

@@

\subsection{IFN reporter cell lines}
We created IFN reporter variants of the A549 human lung epithelial cell line (\FIG{IFNrare}B).
The parental A549 cell line used to create these reporters was obtained from ATCC (CCL-185), and was tested as negative for mycoplasma contamination by the Fred Hutch Genomics Core and authenticated using the ATCC STR profiling service.
The cells were maintained in D10 media (DMEM supplemented with 10\% heat-inactivated fetal bovine serum, 2 mM L-glutamine, 100 U of penicillin / ml, and 100 $\mu$g of streptomycin / ml) at 37$^{\circ}$C and 5\% carbon dioxide.

To create the type I interferon reporters, a 1kb promoeter region upstream of the human IFNB1 gene was cloned into the pHAGE2 lentiviral vector~\citep{oconnell2010lentiviral}, with a NotI site immediately downstream of the promoter serving as an artificial Kozak sequence. 
Downstream of this NotI site, each of the following reporter constructs was cloned: mCherry, mNeonGreen, and mNeonGreen translationally linked to low-affinity nerve growth factor lacking the C-terminal signaling domain (LNGFR$\Delta$C)~\citep{bonini1997hsv,ruggieri1997cell} by a P2A linker~\citep{kim2011high}.
The sequence of the last of these constructs is provided in \FIGDATA[IFNrare]{reporter_sequences}.

To create the type III interferon reporters, a 1.2kb region upstream of the human IL29 (IFNL1) gene was cloned into the pHAGE2 vector, with the native Kozak sequence retained at the 3' end. 
Downstream of this promoter we cloned ZsGreen translationally linked to LNGFR$\Delta$C via a P2A linker.
The sequence of this construct is provided in \FIGDATA[IFNrare]{reporter_sequences}.

We used these constructs to generate lentiviral vectors~\citep{oconnell2010lentiviral} which were used to transduce of A549 cells in the presence of 5 $\mu$g polybrene.
We then sorted single transduced cells and expanded them.
A portion of the expanded cells were tested for reporter activity by transfecting poly(I:C) (a potent agonist of the RIG-I pathway), and we retained clones with strong activation.
Importantly, the cells that we retained for further use were not the same portion that were tested by poly(I:C) treatment, but rather a separate split of the same population---this avoids any selection on the cells from transient activation of IFN.
For the dual type I / type III reporter used in \FIGSUPP[IFNrare]{type_I_vs_III}, a single-cell clone of the type III reporter cell line was transduced with the type I reporter bearing the mCherry fluorescent marker, and then isolated and propagated as a single cell clone for the other cell lines.

\FIGSUPP[IFNrare]{reporter_validation} shows validation of the reporter cell lines using infection with saturating amounts of the Cantell strain of Sendai virus (obtained from Charles River Laboratories).
For detection of the cell-surface bound LNGFR$\Delta$C, \jdbcomment{cells were stained with?}.

\subsection{Viruses for single-cell experiments}
\jdbcomment{Alistair writes this}
All viruses were generated using reverse-genetics as previously described.
All infections were performed in influenza growth media as previously described.
Briefly, bidirectional plasmids producing viral mRNA and vRNA are transfected into a co-culture of MDCK-SIAT1 and 293t cells grown in D10.
After 24h, media is changed from D10 to influenza growth media.
Between 48h and 72h post-infection, depending on viral growth kinetics, supernatant was harvested and titered by TCID50.
Virus was subsequently expanded on MDCK-SIAT1 cells as described below.
For all viral variants, the described mutations were introduced into reverse-genetics plasmids using PCR mutagenesis. 
For the HA-flanked mCherry pseudovirus, a plasmid was generated similarly to a previously-described HA-flanked eGFP vector.
Synonymously-barcoded virus utilized previously described 3' synonymous mutations, and two newly-introduced single-nucleotide synonymous mutations at the 5' end of each viral gene.
Wild-type, synonymously-barcoded, NS1-inactivated, and single-nucleotide variant viruses were all grown in unmodified MDCK-SIAT1 and 293t cells.
Viral variants bearing large internal deleltions in the PB1 segment were grown in MDCK-SIAT1 and 293t cells expressing PB1 as previously described.
Pseudovirus expressing mCherry in place of eGFP was grown in MDCK-SIAT1 and 293t cells expressing HA as previously described.

@@

\subsection{Viruses for validation experiments}
\jdbcomment{Alistair writes this}
All viruses were generated using reverse-genetics as previously described.
All infections were performed in influenza growth media as previously described.
Briefly, bidirectional plasmids producing viral mRNA and vRNA are transfected into a co-culture of MDCK-SIAT1 and 293t cells grown in D10.
After 24h, media is changed from D10 to influenza growth media.
Between 48h and 72h post-infection, depending on viral growth kinetics, supernatant was harvested and titered by TCID50.
Virus was subsequently expanded on MDCK-SIAT1 cells as described below.
For all viral variants, the described mutations were introduced into reverse-genetics plasmids using PCR mutagenesis. 
For the HA-flanked mCherry pseudovirus, a plasmid was generated similarly to a previously-described HA-flanked eGFP vector.
Synonymously-barcoded virus utilized previously described 3' synonymous mutations, and two newly-introduced single-nucleotide synonymous mutations at the 5' end of each viral gene.
Wild-type, synonymously-barcoded, NS1-inactivated, and single-nucleotide variant viruses were all grown in unmodified MDCK-SIAT1 and 293t cells.
Viral variants bearing large internal deleltions in the PB1 segment were grown in MDCK-SIAT1 and 293t cells expressing PB1 as previously described.
Pseudovirus expressing mCherry in place of eGFP was grown in MDCK-SIAT1 and 293t cells expressing HA as previously described.
